---
title: "Lookup tables for Maize data"
author: "Gustavo V. Barroso"
date: "`r Sys.Date()`"
output:
  pdf_document:
  toc: true
number_sections: true
toc_depth: 1
---
  
```{r setup, include=F, message=F, warning=F}

knitr::opts_chunk$set(echo=TRUE)

library(tidyverse)
library(scales)
library(bigsnpr) # for seq_log
library(data.table)
library(cowplot)

knitr::opts_knit$set(root.dir="~/Data/bgs_maize/test_expansion")
```

# Creating folders and files

Builds a parameter grid that allows for fast assembly of whole chromosomes

```{r, params, include=T, eval=T, results=F, message=F, warning=F}
# N's and T's specified from present to past, ommiting last epoch requested by Paulo
# scaled down 10x to handle Ne*s parameters

Ns <- round(seq_log(1.5e+4, 1e+7, 11), digits=0)
Ts <- paste(0, 15, 40, 400, 1000, sep=";")

# scaling u by 10x (Paulo suggested 3.3e-8) due to high order of 1-2p
# this effectively "compensates" the scaling of Ns and Ts above, but will still need to be handled by
# bgshr since the scaling argument there will operate on u, r and s simultaneously
uL <- 1e-6
uR <- 1e-6

# r and s specified as they would normally be for momentspp (ie, not scaling to compensate demo)
r <- c(1e-8, 1e-6, 1e-4, 1e-2)
s <- c(-1e-6, -1e-5)

params <- setDT(crossing(Ns, Ts, r, s, uL, uR))

params$Ns <- paste(params$Ns, 1.5e+4, 1.5e+5, 1.5e+3, 1.5e+4, sep=";")
n_models <- nrow(params)

fwrite(as.data.frame(r), "r.csv", col.names=F)
fwrite(as.data.frame(s), "s.csv", col.names=F)
fwrite(params, "params.csv")
```

Creates directories where files will be written to

```{bash dirs, engine.opts='-i', include=F, eval=T, results=F, warning=F}

# NOTE: this chunk requires that the directories do not already exist
source ~/.bashrc
  
cat r.csv | while IFS=, read -r r
do
  mkdir r_$r
  cat s.csv | while IFS=, read -r s
  do
    mkdir r_$r/s_$s
  done
done

find -mindepth 2 -type d -exec cp opt.bpp {} \;

# makes directory to store purely neutral models to get pi0 predictions
mkdir demo
cp opt.bpp demo
```

Creates files that will be used as input for moments++

```{r, demes, include=F, eval=T, results=F, message=F, warning=F}

params <- fread("params.csv")
n_models <- nrow(params)

# writes augmented demes files specifying models (one demography, r X s)
options(digits=16) # precision
for(i in 1:n_models) {
  
  if(i %% 10000 == 0) {
    print(Sys.time())
    cat(paste(i, "\n"))
  }
  
  Ns <- as.numeric(strsplit(params[i,]$Ns, ";")[[1]])
  Ts <- as.integer(strsplit(params[i,]$Ts, ";")[[1]])
  u <- as.numeric(params[i,]$uL)
  r <- as.numeric(params[i,]$r)
  s <- as.numeric(params[i,]$s)
  
  name <- paste("model_", i, sep="")
  sink(paste("r_", r, "/s_", s,"/", name, ".yaml", sep=""))
  
  cat("time_units: generations\n")
  cat("demes:\n")
  cat("  - name: X\n")
  cat("    epochs:\n")
  for(j in length(Ts):1) {
    cat("      - end_time: ")
    cat(format(Ts[j], scientific=F))
    cat("\n")
    cat("        start_size: ")
    cat(format(Ns[j], scientific=F))
    cat("\n")
  }
  cat("metadata:\n")
  cat("  - name: mutation\n")
  cat("    left_factor: 1\n")
  cat("    epochs:\n")
  cat("      - end_time: 0\n")
  cat("        rates: [")
  cat(u)
  cat("]\n")
  cat("  - name: recombination\n")
  cat("    epochs:\n")
  cat("      - end_time: 0\n")
  cat("        rates: [")
  cat(r)
  cat("]\n")
  cat("  - name: selection\n")
  cat("    start_time: .inf\n")
  cat("    epochs:\n")
  cat("      - end_time: 0\n")
  cat("        rates: [")
  cat(format(s, scientific=F))
  cat("]\n")
  
  sink()
}

# creates the (purely neutral) demographic model
# single u* and r because we're only interested in pi0 here (for B-values)
Ns <- as.numeric(strsplit(params[i,]$Ns, ";")[[1]])
Ts <- as.numeric(strsplit(params[i,]$Ts, ";")[[1]])
u <- as.numeric(unique(params$uL))

sink(paste("demo/model_neutral.yaml", sep=""))
cat("time_units: generations\n")
cat("demes:\n")
cat("  - name: X\n")
cat("    epochs:\n")
for(j in length(Ts):1) {
  cat("      - end_time: ")
  cat(Ts[j])
  cat("\n")
  cat("        start_size: ")
  cat(Ns[j])
  cat("\n")
}
cat("metadata:\n")
cat("  - name: mutation\n")
cat("    left_factor: 1\n")
cat("    epochs:\n")
cat("      - end_time: 0\n")
cat("        rates: [")
cat(u)
cat("]\n")
cat("  - name: recombination\n")
cat("    epochs:\n")
cat("      - end_time: 0\n")
cat("        rates: [0]\n")
cat("  - name: selection\n")
cat("    start_time: .inf\n")
cat("    epochs:\n")
cat("      - end_time: 0\n")
cat("        rates: [0]\n")
sink()
```

# Running moments++

Computes summary statistics for each model
For reference, the following script was run, not from this Rmd file but looping from the terminal

```{bash moments++, engine.opts='-i', include=F, eval=F, results=F, warning=F}
#!/bin/bash
f=$(find -name '*.yaml' | cut -d '/' -f 2)

N=150000 # largest N in prescribed (scaled down) demography
x=$(sed '23q;d' $f | cut -d : -f 2 | cut -d [ -f 2 | cut -d ] -f 1)
s=$(echo "scale=10; $x" | bc)
a=$(echo "scale=10; $N * $s" | bc)
if [[ $(echo "$a < -1" | bc) == 1 ]]
then
  b=${a%.*}
  Ns=$(($b * -2))
  if [[ $Ns -lt 20 ]]
  then
    momentspp param=opt.bpp F=$f O=60 > /dev/null
  elif [[ $Ns -lt 70 ]]
  then
    momentspp param=opt.bpp F=$f O=150 > /dev/null
  elif [[ $Ns -lt 150 ]]
  then
    momentspp param=opt.bpp F=$f O=300 > /dev/null
  else
    momentspp param=opt.bpp F=$f O=400 > /dev/null
  fi
else
  momentspp param=opt.bpp F=$f O=10 > /dev/null
fi
```

Collects lookup table with Hr values on the parameter grid

```{r, lookup, include=T, eval=F, results=F, message=F, warning=F}

params <- fread("params.csv")
n_models <- nrow(params)

lookup_r <- unique(params$r)
lookup_s <- unique(params$s)
t <- max(as.numeric(strsplit(as.character(unique(params$Ts)), ";")[[1]]))
sampling_times <- seq(from=0, to=0, by=-500) # no temporal sampling

hrs <- as.data.frame(matrix(nrow=n_models, ncol=length(sampling_times)))
hls <- as.data.frame(matrix(nrow=n_models, ncol=length(sampling_times)))

names(hrs) <- as.character(sampling_times)
names(hls) <- as.character(sampling_times)

hrs$Order <- 0
hls$Order <- 0

for(i in 1:length(lookup_r)) {
  cat(paste(i, "\n"))
  print(Sys.time())
  for(j in 1:length(lookup_s)) {
    fnames <- list.files(paste("r_", lookup_r[i], "/s_", lookup_s[j], sep=""),
                         pattern="*_expectations.txt", full.names=TRUE)
    
    idx <- 0  
    for(l in 1:length(fnames)) {
      moms <- fread(fnames[l], header=F)
      str_sides <- strsplit(fnames[l], "_O_")[[1]]
      idx <- as.numeric(strsplit(str_sides[1], "model_")[[1]][2])
      order <- as.numeric(strsplit(str_sides[2], "_e")[[1]][1])
      
      left <- filter(moms, V1=="Hl_0_0")
      right <- filter(moms, V1=="Hr_0_0")
      
      hls[idx,] <- c(t(dplyr::select(left, V3)), order)
      hrs[idx,] <- c(t(dplyr::select(right, V3)), order)
    }  
  }
}

lookup_hl <- cbind.data.frame(params, hls)
lookup_hr <- cbind.data.frame(params, hrs)
lookup_tbl <- pivot_longer(lookup_hr, cols=as.character(sampling_times),
                           names_to="Generation", values_to="Hr")
lookup_tbl$Hr <- as.numeric(lookup_tbl$Hr)
lookup_tbl_hl <- pivot_longer(lookup_hl, cols=as.character(sampling_times),
                              names_to="Generation", values_to="Hl")
lookup_tbl$Hl <- as.numeric(lookup_tbl_hl$Hl)
lookup_tbl$Generation <- as.numeric(lookup_tbl$Generation)

# get pi0 from demography
#demo_models <- crossing(unique(params$Ns), unique(params$Ts))
demo_models <- cbind.data.frame(unique(params$Ns), pi0)
#demo_models$model <- 1:nrow(demo_models)
names(demo_models) <- c("Ns", "Ts")
num_models <- nrow(demo_models)
pi0 <- as.data.frame(matrix(nrow=num_models, ncol=length(sampling_times)))
names(pi0) <- as.character(sampling_times)

#fnames <- list.files("demo", pattern="*_e_*.*_time.txt", full.names=TRUE) if time_steps > 0
fnames <- list.files("demo", pattern="*_expectations.txt", full.names=TRUE)
for(l in 1:length(fnames)) {
  moms <- fread(fnames[l])
  str_sides <- strsplit(fnames[l], "_O_")[[1]]
  order <- as.numeric(strsplit(str_sides[2], "_e")[[1]][1])
  idx <- as.numeric(strsplit(str_sides[1], "model_")[[1]][2])
  right <- filter(moms, V1=="Hr_0_0")
  pi0[idx, 1] <- t(dplyr::select(filter(right, V1=="Hr_0_0"), V3))
}


demo_pi0 <- cbind.data.frame(demo_models, pi0)
m_pi0 <- pivot_longer(demo_pi0, 
                      cols=as.character(sampling_times),
                      names_to="Generation", values_to="pi0")
m_pi0$Generation <- as.numeric(m_pi0$Generation)
m_pi0 <- dplyr::select(m_pi0, -c(Ts))

lookup_tbl <- left_join(lookup_tbl, m_pi0, by=c("Ns", "Generation"))
lookup_tbl$B <- lookup_tbl$Hr / lookup_tbl$pi0

# when in ne_bar_ss, equivalent generation in temporal model
#lookup_tbl$Generation <- rep(sampling_times, nrow(lookup_tbl) / length(sampling_times)) 

fwrite(lookup_tbl, "lookup_tbl.csv.gz", compress="gzip") 
```


```{r, plots, eval=F}

tbl <- lookup_tbl
tbl$exp_fold <- as.factor(round(as.numeric(unlist(lapply(strsplit(tbl$Ns, ";"),
                                   function(y) return(y[1])))) / 15000, 0))

p <- ggplot(data=tbl, aes(y=B^1000, x=r/10, color=exp_fold)) + 
  geom_point() + geom_line() + theme_bw() +
  facet_wrap(~s/10) + scale_x_log10(breaks=unique(tbl$r/10)) +
  labs(title=NULL,
           x="r", y="B", color="Expansion Factor") +
  theme(strip.text.y=element_text(size=16),      
        strip.text.x=element_text(size=16),       
        axis.title=element_text(size=16), 
        axis.text=element_text(size=12), 
        axis.text.x=element_text(size=12),
        legend.text=element_text(size=16),
        legend.title=element_text(size=16),
        legend.position="bottom")

save_plot("test_expansion.pdf", p, base_height=8, base_width=16)
```
