---
title: "BGS under non-equilibrium demography (paper 1)"
author: "Gustavo V. Barroso"
date: "`r Sys.Date()`"
output:
  pdf_document:
  toc: true
number_sections: true
toc_depth: 1
---
  
```{r setup, include=F, message=F, warning=F}

knitr::opts_chunk$set(echo=TRUE)

library(tidyverse)
library(bigsnpr) # for seq_log
library(data.table)

library(reticulate)
use_python("/usr/bin/python3")

knitr::opts_knit$set(root.dir="~/Data/momentspp/paper_1/equilibrium/")
```

This is an empty python chunk.
Apparently this must exist in order for reticulate to be able to read
R data frames in future python chunks.

```{python}
```

# Creating folders and files

This builds a look-up table on a dense parameter grid that will allow for fast 
assembly of whole chromosomes

```{r, params, include=F, eval=T, results=F, message=F, warning=F}

Na <- 1e+4 # ancestral N
N1 <- c(Na / 10, Na * 10) # current N
t <- 2 * Na # in generations

options(digits=16)

# rounding is required so that directory names don't diverge due to precision
uL <- c(1e-8)
r <- c(1e-8, 2.5e-05, 1e-03)
s <- c(-1e-3, -1e-4, -1e-5)

# for discretizing Gamma distributions
dt_u <- data.table(uL)
dt_r <- data.table(r)
dt_s <- data.table(s)

setkey(dt_u, uL)
setkey(dt_r, r)
setkey(dt_s, s)

params <- setDT(crossing(Na, N1, t, r, s, uL))
n_models <- nrow(params)

fwrite(dt_u, "u.csv", col.names=F)
fwrite(dt_r, "r.csv", col.names=F)
fwrite(dt_s, "s.csv", col.names=F)
fwrite(params, "params.csv")
```

```{bash dirs, engine.opts='-i', include=F, eval=T, results=F, warning=F}

# NOTE: this chunk requires that the directories do not already exist
source ~/.bashrc
  
cat r.csv | while IFS=, read -r r
do
  mkdir r_$r
  cat s.csv | while IFS=, read -r s
  do
    mkdir r_$r/s_$s
    cat u.csv | while IFS=, read -r u
    do
      mkdir r_$r/s_$s/u_$u
    done
  done
done

find -mindepth 3 -type d -exec cp opt.bpp {} \;

# makes directory to store purely neutral models to get pi0 predictions
mkdir demo
cp opt.bpp demo
```

Creates files that will be used as input for moments++

```{r, demes, include=F, eval=T, results=F, message=F, warning=F}

params <- fread("params.csv")

options(digits=16)
# writes demes files specifying models
for(i in 1:n_models) {
  
  if(i %% 10000 == 0) {
    print(Sys.time())
    cat(paste(i, "\n"))
  }
  
  Na <- as.integer(params[i,]$Na)
  N1 <- as.integer(params[i,]$N1)
  t <- as.integer(params[i,]$t)
  u <- as.numeric(params[i,]$uL)
  r <- as.numeric(params[i,]$r)
  s <- as.numeric(params[i,]$s)
  
  name <- paste("model_", i, sep="")
  sink(paste("r_", r, 
             "/s_", s,
             "/u_", u,
             "/",
    name, ".yaml", sep=""))
  
  cat("time_units: generations\n")
  cat("demes:\n")
  cat("  - name: X\n")
  cat("    epochs:\n")
  if(N1 != Na) {
    cat("      - end_time: ")
    cat(t)
    cat("\n")
    cat("        start_size: ")
    cat(Na)
    cat("\n")
  }
  cat("      - end_time: 0\n")
  cat("        start_size: ")
  cat(Na)
  cat("\n")
  cat("metadata:\n")
  cat("  - name: mutation\n")
  cat("    left_factor: 1\n")
  cat("    epochs:\n")
  cat("      - end_time: 0\n")
  cat("        rates: [")
  cat(u)
  cat("]\n")
  cat("  - name: recombination\n")
  cat("    epochs:\n")
  cat("      - end_time: 0\n")
  cat("        rates: [")
  cat(r)
  cat("]\n")
  cat("  - name: selection\n")
  cat("    start_time: .inf\n")
  cat("    epochs:\n")
  cat("      - end_time: 0\n")
  cat("        rates: [")
  cat(format(s, scientific=F))
  cat("]\n")
  
  sink()
}

# creates (purely neutral) demographic models
Na <- as.integer(unique(params$Na))
N1 <- as.integer(unique(params$N1))
t <- as.integer(unique(params$t))
u <- as.numeric(unique(params$uL))

c <- 1
for(N in N1) {
  for(mu in u) {
    sink(paste("demo/model_", c, ".yaml", sep=""))
    cat("time_units: generations\n")
    cat("demes:\n")
    cat("  - name: X\n")
    cat("    epochs:\n")
    if(N != Na) {
      cat("      - end_time: ")
      cat(t)
      cat("\n")
      cat("        start_size: ")
      cat(Na)
      cat("\n")
    }
    cat("      - end_time: 0\n")
    cat("        start_size: ")
    cat(N)
    cat("\n")
    cat("metadata:\n")
    cat("  - name: mutation\n")
    cat("    left_factor: 1\n")
    cat("    epochs:\n")
    cat("      - end_time: 0\n")
    cat("        rates: [")
    cat(mu)
    cat("]\n")
    cat("  - name: recombination\n")
    cat("    epochs:\n")
    cat("      - end_time: 0\n")
    cat("        rates: [0]\n")
    cat("  - name: selection\n")
    cat("    start_time: .inf\n")
    cat("    epochs:\n")
    cat("      - end_time: 0\n")
    cat("        rates: [0]\n")
    sink()
    
    c <- c + 1
  }
}
  
```

# Running moments++

Computes summary statistics for each parameter combination in lookup_tbl

```{bash moments++, engine.opts='-i', include=F, eval=F, results=F, warning=F}

#!/bin/bash

i=1
nr=$(wc -l r.csv | cut -d ' ' -f 1)

for r in r_*
do
  date
  echo "$i of $nr"
  cd $r
  for s in s_*
  do
    cd $s
    for u in u_*
    do
      cd $u
      for f in model_*.yaml
      do
	      Na=$(sed '6q;d' $f | cut -d : -f 2 | cut -d ' ' -f 2)
	      N1=$(sed '8q;d' $f | cut -d : -f 2 | cut -d ' ' -f 2)
	      Ne=$Na
	      if [[ $N1 -gt $Na ]]
        then
	        Ne=$N1
	      fi
        x=$(sed '23q;d' $f | cut -d : -f 2 | cut -d [ -f 2 | cut -d ] -f 1)
        s=$(echo "scale=10; $x" | bc)
	      a=$(echo "scale=10; $Ne * $s" | bc)
        if [[ $(echo "$a < -1" | bc) == 1 ]]
	      then
	        b=${a%.*}
	        Ns=$(($b * -1))
	        if [[ $Ns -lt 10 ]]
	        then
  	        momentspp params=opt.bpp F=$f O=50 > /dev/null
  	      elif [[ $Ns -lt 50 ]]
  	      then 
  	        momentspp params=opt.bpp F=$f O=200 > /dev/null
  	      else
  	        momentspp params=opt.bpp F=$f O=300 > /dev/null
  	      fi
        else
	        momentspp params=opt.bpp F=$f O=10 > /dev/null
	      fi
      done
      cd ..
    done
    cd ..
  done
  cd ..
  i=$(($i + 1))
done

# get pi0 over time
cd demo
for f in *yaml
do 
  momentspp params=opt.bpp F=$f O=2
done
```

Collects lookup table with Hr values on the parameter grid

```{r, grid, include=F, eval=F, results=F, message=F, warning=F}

params <- fread("params.csv")

lookup_r <- unique(params$r)
lookup_s <- unique(params$s)
lookup_u <- unique(params$uL)

n_models <- nrow(params)
t <- unique(params$t)
sampling_times <- seq(from=0, to=t, by=10)

# Het stats over time (include steady-state and present-time)
hrs <- as.data.frame(matrix(nrow=n_models, ncol=1))#length(sampling_times)))
hls <- as.data.frame(matrix(nrow=n_models, ncol=1))#length(sampling_times)))

names(hrs) <- as.character(sampling_times)
names(hls) <- as.character(sampling_times)

hrs$Order <- 0
hls$Order <- 0

for(i in 1:length(lookup_r)) {
  cat(paste(i, "\n"))
  print(Sys.time())
  for(j in 1:length(lookup_s)) {
    for(k in 1:length(lookup_u)) {
      fnames <- list.files(paste("r_", lookup_r[i], "/s_", lookup_s[j], 
                                 "/u_", lookup_u[k], sep=""),
                           pattern="*_e_1_expectations.txt", full.names=TRUE)
      
      for(l in 1:length(fnames)) {
        moms <- fread(fnames[l])
        str_sides <- strsplit(fnames[l], "_O_")[[1]]
        idx <- as.numeric(strsplit(str_sides[1], "model_")[[1]][2])
        order <- as.numeric(strsplit(str_sides[2], "_expectations.txt")[[1]][1])
        
        hls[idx,] <- c(t(dplyr::select(filter(moms, V1=="Hl_0_0"), V3)), order)
        hrs[idx,] <- c(t(dplyr::select(filter(moms, V1=="Hr_0_0"), V3)), order)
      }
    }
  }
}

params_hl <- cbind.data.frame(params, hls)
params_hr <- cbind.data.frame(params, hrs)

fwrite(params_hl, "hl_time.csv.gz", compress="gzip") # used by other scripts
fwrite(params_hr, "hr_time.csv.gz", compress="gzip") # used by other scripts
```
