---
title: "Benchmarking Moments++ in the 1-population case"
author: "Gustavo V. Barroso"
date: "`r Sys.Date()`"
output:
  pdf_document:
  toc: true
number_sections: true
toc_depth: 3
---
  
```{r setup, include=F, message=F, warning=F}

knitr::opts_chunk$set(echo=TRUE)

library(tidyverse)
library(data.table)
library(cowplot)
library(stringr)
library(RColorBrewer)
library(scales)
library(plotly)
library(MASS)
library(lmtest)
library(nlme)
library(car)
library(bigsnpr)
library(data.table)
library(GenomicRanges)
library(pracma)

library(reticulate)
use_python("/usr/bin/python3")

scale.4d <- function(x) sprintf("%.4f", x)

knitr::opts_knit$set(root.dir="~/Data/momentspp/")
```

This is an empty python chunk.
Apparently this must exist in order for reticulate to be able to read
R data frames in future python chunks.

```{python}

```

Here we benchmark against moments.TwoLocus using equilibrium demography
to assess accuracy of moments++ under strong selection (adding (1-2p)^k factors)

### Building Demes files

```{r, demes-selected, include=F, eval=T, results=F, message=F, warning=F}

Ns <- 20000
Ts <- 0
u <- 1e-8
r  <- c(1e-3, 1e-4, 1e-5, 1e-6)
s <- c(-1e-5, -1e-04, -1e-3, -2e-3, -5e-3)

params <- crossing(Ns, u, r, s)
params$scenario <- 1:nrow(params)
n_models <- nrow(params)

fwrite(params, "params.csv")

Sys.setenv(num_models=n_models)

py$params <- params # stores df to be used in python sections
py$n_models <- as.integer(n_models + 1) # because python is indexed from zero

for(i in 1:n_models) {
  
  name <- paste("model_", i, sep="")
  sink(paste(name, ".yaml", sep=""))
  
  cat("time_units: generations\n")
  cat("demes:\n")
  cat("  - name: X\n")
  cat("    epochs:\n")
  cat("      - end_time: 0\n")
  cat("        start_size: ")
  cat(params$Ns[i])
  cat("\n")
  cat("metadata:\n")
  cat("  - name: mutation\n")
  cat("    epochs:\n")
  cat("      - end_time: 0\n")
  cat("        rates: [")
  cat(params$u[i])
  cat("]\n")
  cat("  - name: recombination\n")
  cat("    epochs:\n")
  cat("      - end_time: 0\n")
  cat("        rates: [")
  cat(params$r[i])
  cat("]\n")
  cat("  - name: selection\n")
  cat("    start_time: .inf\n")
  cat("    epochs:\n")
  cat("      - end_time: 0\n")
  cat("        rates: [")
  cat(params$s[i])
  cat("]\n")
  
  sink()
}

```

### Running moments++

```{bash run-moments++-, engine.opts='-i', include=F, eval=T}

source ~/.bashrc

for((i=1;i<=$num_models;i++))
do
  momentspp param=opt.bpp F=model_$i.yaml O=25;
  momentspp param=opt.bpp F=model_$i.yaml O=50;
  momentspp param=opt.bpp F=model_$i.yaml O=75;
  momentspp param=opt.bpp F=model_$i.yaml O=100;
  momentspp param=opt.bpp F=model_$i.yaml O=150;
  momentspp param=opt.bpp F=model_$i.yaml O=200;
  momentspp param=opt.bpp F=model_$i.yaml O=300;
done
```

### Running moments.TwoLocus

For benchmarking the single population case, we use equilibrium demography:
  
```{python run-momentsTwoLocus, include=F, eval=T, results=F}

import moments.TwoLocus
import sys, os
import numpy as np

for i in range(1, n_models):
  with open(f"moments_twoLocus_n40_model_{i}.csv", "w+") as fout:
      n_samples = 40
      
      # i - 1 because python is indexed from zero
      Ne = params["Ns"][i - 1] 
      u = params["u"][i - 1] 
      r = params["r"][i - 1] 
      s = params["s"][i - 1]
      
      gammas = [2 * Ne * s, 2 * Ne * s, 0] # sel on [AB, Ab, aB] haplotypes, resp
      rho = 4 * Ne * r
      theta = 4 * Ne * u
      
      F = moments.TwoLocus.Demographics.equilibrium(n_samples, 
                                                    rho=rho, 
                                                    theta=theta, 
                                                    sel_params=gammas)
      
      fs = moments.Spectrum(moments.LinearSystem_1D.steady_state_1D(n_samples,
                           gamma=2 * Ne * s) * theta)    
      fs_left = moments.Spectrum(F[0, :, 0])
      fs_right = moments.Spectrum(F[0, 0, :])
      
      fout.write(f"D^2 = {F.D2():}\n")
      fout.write(f"Dz = {F.Dz():}\n")
      fout.write(f"Hl = {fs_left.pi()}\n")
      fout.write(f"Hr = {fs_right.pi()}\n")
      fout.write(f"pi2 = {F.pi2():}\n")
      
      assert np.allclose(fs_left, fs)

for i in range(1, n_models):
  with open(f"moments_twoLocus_n80_model_{i}.csv", "w+") as fout:
    n_samples = 80
    
    # i - 1 because python is indexed from zero
    Ne = params["Ns"][i - 1] 
    u = params["u"][i - 1] 
    r = params["r"][i - 1] 
    s = params["s"][i - 1]
    
    gammas = [2 * Ne * s, 2 * Ne * s, 0] # sel on [AB, Ab, aB] haplotypes, resp
    rho = 4 * Ne * r
    theta = 4 * Ne * u
    
    F = moments.TwoLocus.Demographics.equilibrium(n_samples, 
                                                  rho=rho, 
                                                  theta=theta, 
                                                  sel_params=gammas)
    
    fs = moments.Spectrum(moments.LinearSystem_1D.steady_state_1D(n_samples, 
                          gamma=2 * Ne * s) * theta)    
    fs_left = moments.Spectrum(F[0, :, 0])
    fs_right = moments.Spectrum(F[0, 0, :])
    
    fout.write(f"D^2 = {F.D2():}\n")
    fout.write(f"Dz = {F.Dz():}\n")
    fout.write(f"Hl = {fs_left.pi()}\n")
    fout.write(f"Hr = {fs_right.pi()}\n")
    fout.write(f"pi2 = {F.pi2():}\n")
    
    assert np.allclose(fs_left, fs)
```

### Plots

```{r, plots-selected-1-pop, include=T, eval=T, results=F, message=F, warning=F}

params <- fread("params.csv")
n_models <- nrow(params)

commom_stats <- c("DD_0_0",
                  "Dr_0_0_(1-2p0)^1",
                  "Hl_0_0", 
                  "Hr_0_0",
                  "pi2_0_0_0_0")

# from moments++
orders <- c(25, 50, 75, 100, 150, 200, 300)
tbl <- data.frame()
for(j in 1:length(orders)) {
  expectation <- numeric()
  for(i in 1:n_models) {
    vals <- read.table(paste("model_",  i, "_O_", orders[j], 
                             "_expectations.txt", sep=""))
    vals <- subset(vals, V1 %in% commom_stats)$V3
    expectation <- c(expectation, vals)
  }
  
  tmp <- as.data.frame(expectation)
  tmp$stats <- commom_stats
  scenario <- NULL
  
  for(i in 1:n_models) { 
    scenario <- c(scenario, rep(i, length(commom_stats)))
  }
  tmp$scenario <- scenario
  tmp$order <- orders[j]
  
  tbl <- rbind.data.frame(tmp, tbl)
}

mpp <- full_join(tbl, params, by="scenario")
mpp$alpha <- 2 * mpp$s * mpp$N

# moments.TwoLocus was fitted with 2 sample sizes, n==40 and n==80
# n==40
expectation <- numeric()
for(i in 1:n_models) {
  vals <- read.table(paste ("moments_twoLocus_n40_model_", i,".csv", sep=""))$V3
  expectation <- c(expectation, vals)
}

tbl_py_n40 <- as.data.frame(expectation)
tbl_py_n40$stats <- commom_stats
tbl_py_n40$scenario <- scenario
tbl_py_n40$sample_size <- 40
tbl_py_n40 <- full_join(tbl_py_n40, params, by="scenario")

# from 2p(1-p) to p(1-p)
tbl_py_n40[tbl_py_n40$stats=="Hl_0_0", 1] <-
  tbl_py_n40[tbl_py_n40$stats=="Hl_0_0", 1] / 2
tbl_py_n40[tbl_py_n40$stats=="Hr_0_0", 1] <-
  tbl_py_n40[tbl_py_n40$stats=="Hr_0_0", 1] / 2

# n==80
expectation <- numeric()
for(i in 1:n_models) {
  vals <- read.table(paste("moments_twoLocus_n80_model_", i, ".csv", sep=""))$V3
  expectation <- c(expectation, vals)
}

tbl_py_n80 <- as.data.frame(expectation)
tbl_py_n80$stats <- commom_stats
tbl_py_n80$scenario <- scenario
tbl_py_n80$sample_size <- 80
tbl_py_n80 <- full_join(tbl_py_n80, params, by="scenario")

# from 2p(1-p) to p(1-p)
tbl_py_n80[tbl_py_n80$stats=="Hl_0_0", 1] <-
  tbl_py_n80[tbl_py_n80$stats=="Hl_0_0", 1] / 2
tbl_py_n80[tbl_py_n80$stats=="Hr_0_0", 1] <-
  tbl_py_n80[tbl_py_n80$stats=="Hr_0_0", 1] / 2

tbl_py_ns <- bind_rows(tbl_py_n40, tbl_py_n80)

tbl_pp <- full_join(mpp, tbl_py_ns, by=c("stats", colnames(params)))
names(tbl_pp)[1] <- "mpp"
names(tbl_pp)[10] <- "m2l"
tbl_pp <- tbl_pp %>% mutate(ratio=mpp/m2l)

# plots
p0 <- ggplot(data=filter(tbl_pp, stats==commom_stats[3],
                         sample_size==80), aes(x=order, y=ratio)) +
      geom_point(size=3) + geom_line() + theme_bw() + 
  facet_wrap(~alpha, ncol=1) + 
  scale_x_continuous(breaks=unique(tbl_pp$order)) +
  labs(title=NULL, x="Order (1-2p)", y=expression(H[l])) +
  geom_hline(yintercept=1, linetype="dashed", color="red") +
  theme(axis.title=element_text(size=12),
        axis.text=element_text(size=10),
        axis.text.x=element_text(angle=90, size=12, vjust=0.5, hjust=1.0),
        legend.position="bottom")
save_plot("Hl_ratios_order.pdf", p0, base_height=24, base_width=12)

p1 <- ggplot(data=filter(tbl_pp, stats==commom_stats[1],
                         sample_size==80), aes(x=r, y=ratio)) +
      geom_point(size=3) + geom_line() + theme_bw() + 
  facet_grid(+alpha~order) + scale_x_log10(breaks=r) +
  labs(title=NULL, x="r", y=expression(D^2)) +
  geom_hline(yintercept=1, linetype="dashed", color="red") +
  theme(axis.title=element_text(size=12),
        axis.text=element_text(size=10),
        axis.text.x=element_text(angle=90, size=12, vjust=0.5, hjust=1.0),
        legend.position="bottom")
save_plot("dd_ratios.pdf", p1, base_height=24, base_width=12)

p2 <- ggplot(data=filter(tbl_pp, stats==commom_stats[2], 
                         sample_size==80), aes(x=r, y=ratio)) +
      geom_point(size=3) + geom_line() + theme_bw() + 
  facet_grid(+alpha~order) + scale_x_log10(breaks=r) + 
  labs(title=NULL, x="r", y="Dz") +
  geom_hline(yintercept=1, linetype="dashed", color="red") +
  theme(axis.title=element_text(size=12),
        axis.text=element_text(size=10),
        axis.text.x=element_text(angle=90, size=12, vjust=0.5, hjust=1.0),
        legend.position="bottom")
save_plot("dz_ratios.pdf", p2, base_height=24, base_width=12)

p3 <- ggplot(data=filter(tbl_pp, stats==commom_stats[3], 
                         sample_size==80), aes(x=r, y=ratio)) +
      geom_point(size=3) + geom_line() + theme_bw() + 
  facet_grid(+alpha~order) + scale_x_log10(breaks=r) + 
  labs(title=NULL,
       x="r", y=expression(H[l])) +
  geom_hline(yintercept=1, linetype="dashed", color="red") +
  theme(axis.title=element_text(size=12),
        axis.text=element_text(size=10),
        axis.text.x=element_text(angle=90, size=12, vjust=0.5, hjust=1.0),
        legend.position="bottom")
save_plot("hl_ratios.pdf", p3, base_height=24, base_width=12)

y <- ggplot(data=filter(tbl_pp, stats==commom_stats[3], 
                        sample_size==80, alpha < -1),
            aes(x=order, y=ratio)) + 
  geom_point(size=3) + geom_line() + theme_bw() + 
  facet_wrap(~alpha, ncol=1) + 
  scale_x_continuous(breaks=unique(tbl_pp$order)) +
  labs(title=NULL, x="Order (1-2p)", y=expression(H[l])) +
  geom_hline(yintercept=1, linetype="dashed", color="red") +
  theme(axis.title=element_text(size=12),
        axis.text=element_text(size=10),
        axis.text.x=element_text(angle=90, size=12, vjust=0.5, hjust=1.0),
        legend.position="bottom")
save_plot("~/Data/mpp_figures/figures/hl_ratios.pdf", y, 
          base_height=15, base_width=9)

p4 <- ggplot(data=filter(tbl_pp, stats==commom_stats[4],
                         sample_size==80), aes(x=r, y=ratio)) +
      geom_point(size=3) + geom_line() + theme_bw() + 
  facet_grid(+alpha~order) + scale_x_log10(breaks=r) +
  labs(title=NULL,
       x="r", y=expression(H[r])) +
  geom_hline(yintercept=1, linetype="dashed", color="red") +
  theme(axis.title=element_text(size=12),
        axis.text=element_text(size=10),
        axis.text.x=element_text(angle=90, size=12, vjust=0.5, hjust=1.0),
        legend.position="bottom")
save_plot("hr_ratios.pdf", p4, base_height=24, base_width=12)

p5 <- ggplot(data=filter(tbl_pp, stats==commom_stats[5], 
                         sample_size==80), aes(x=r, y=ratio)) +
      geom_point(size=3) + geom_line() + theme_bw() +  
  facet_grid(+alpha~order) + scale_x_log10(breaks=r) + 
  labs(title="Ratios mpp / m2l, rows->alpha, cols->order",
       x="r", y=expression(pi[2])) +
  geom_hline(yintercept=1, linetype="dashed", color="red") +
  theme(axis.title=element_text(size=12),
        axis.text=element_text(size=10),
        axis.text.x=element_text(angle=90, size=12, vjust=0.5, hjust=1.0),
        legend.position="bottom")
save_plot("pi2_ratios.pdf", p5, base_height=24, base_width=12)

# compare to HK95 pred
tbl_hr <- filter(tbl_pp, stats=="Hr_0_0", sample_size==80)
tbl_hr$hk95 <- 2 * tbl_hr$N * tbl_hr$u * 
    (1 - (tbl_hr$u * -1 * tbl_hr$s) /  (2 * (-tbl_hr$s + tbl_hr$r)^2))
molten_hr <- pivot_longer(tbl_hr, cols=c(mpp, hk95), names_to="method")

for(n in unique(molten_hr$N)) {
  q <- ggplot(data=filter(molten_hr, N==n, r==1e-6), 
              aes(x=alpha, y=value, color=method)) +
        geom_point(size=3) + geom_line() + theme_bw() + facet_wrap(~order) +
    scale_x_continuous(breaks=pretty_breaks()) +
    scale_y_continuous(breaks=pretty_breaks()) +
    labs(title=paste("Hr, N=", n, ", r=1e-8", sep=""),
         x=expression(alpha), y=expression(H[r])) +
    theme(axis.title=element_text(size=12),
          axis.text=element_text(size=10),
          axis.text.x=element_text(angle=90, size=12, vjust=0.5, hjust=1.0),
          legend.position="bottom")
  q
  save_plot(paste("Hr, N=", n, ", r=1e-6.pdf", sep=""), q,
            base_height=10, base_width=12)
}

for(a in unique(molten_hr$alpha)) {
 r <- ggplot(data=filter(molten_hr, alpha==a), aes(x=r, y=value, color=method)) +
      geom_point(size=3) + geom_line() + theme_bw() + facet_wrap(~order) +
  scale_x_log10() +
  scale_y_continuous(breaks=pretty_breaks()) +
  labs(title=paste("Hr, alpha=", a, sep=""), x="r", y=expression(H[r])) +
  theme(axis.title=element_text(size=12),
        axis.text=element_text(size=10),
        axis.text.x=element_text(angle=90, size=12, vjust=0.5, hjust=1.0),
        legend.position="bottom")
  r
  save_plot(paste("Hr, alpha=", a, ".pdf", sep=""), r,
            base_height=10, base_width=12)
}
```
